The unequal distribution of access to education and transport: insights from

### 2016-06-06 paper.Rmd
Add back Ana's content

The SPMR presents an excellent opportunity to measure the extent of educational and transport infrastructure, and to study the impact of public policies on different socio-economic groups. Partly as a response to the pressures from the Free Pass Movement that started in March, 2013, low income public school students residing in the SPMR gained access to a public transport subsidy in 2015. However, in October 2015, the São Paulo state government announced during a television interview that as part of a budgetary deficit reduction plan, dozens of secondary schools would be closed in 2016. The change in policy was estimated to affect over 300,000 students, many whom would be placed in schools far from their homes. As a response, in the following months students occupied over 200 schools. Protesters scored some victories: the reform was postponed for one year, and the education secretary resigned (The New York Times, 2015).

Although this time the policy change was not carried out, in a context of high spatial inequalities in the provision of public transport and schooling, it is worth asking whether a public transport subsidy for students can compensate for the lower provision of public schools in some areas. In this dynamic and highly political context, new quantitative evidence can help shed light on the the relationship between transport and educational inequalities and the extent to which they are mutually reinforcing. Moreover, the results should help design policies that are more inclusive and equitable by highlighting areas where the provision of schools, travel to school opportunities, or both, are particularly restrictive.

@@

Our last source of information is an Google Distance Matrix API. We use this API to find the travel times and distances by public transport. This was implemented using an interface to the API by means of the R package **stplanr** [@stplanr]. To provide an example of how the method works, a reproducible example is illustrated below.

>>>>

The SPMR presents an excellent opportunity to measure the extent of educational and transport infrastructure, and to study the impact of public policies on different socio-economic groups. Partly as a response to the pressures from the Free Pass Movement that started in March, 2013, low income public school students residing in the SPMR gained access to a public transport subsidy in 2015. However, in October 2015, the São Paulo state government announced during a television interview that as part of a budgetary deficit reduction plan, dozens of secondary schools would be closed in 2016. The change in policy was estimated to affect more than 300,000 students, many whom would be placed in schools far from their homes. As a response, in the following months students occupied over 200 schools. Protesters scored some victories: the reform was postponed for one year, and the education secretary resigned (The New York Times, 2015).

Although this time the policy change was not carried out, in a context of high spatial inequalities in the provision of public transport and schooling, it is worth asking whether a public transport subsidy for students can compensate for the lower provision of public schools in some areas. In this dynamic and highly political context, new quantitative evidence can help shed light on the relationship between transport and educational inequalities and the extent to which they are mutually reinforcing. Moreover, the results should help design policies that are more inclusive and equitable by highlighting areas where the provision of schools, travel to school opportunities, or both, are particularly restrictive.

@@

The final source of information was derived from the Google Distance Matrix API, which provided travel times and distances by public transport. This was implemented using an interface to the API by means of the R package **stplanr** [@stplanr]. To provide an example of how the method works, a reproducible example is illustrated below.
